{\bf {\Large Improving estimates of ecosystem metabolism computed from dissolved oxygen time series}}

### 2014-11-24 abstract.tex
revisions that include sun and tide angle diagnostics

In aquatic ecosystems, time series of \ac{DO} have been used to compute estimates of integrated ecosystem metabolism.  Central to this open water or ``Odum'' method is the assumption that the dissolved oxygen time series is a Lagrangian specification of the flow field.  However, most \ac{DO} time series are collected at fixed locations, such the method must assume that changes in dissolved oxygen principally reflect ecosystem metabolism and that effects due to advection or mixing can be neglected.  A statistical model using weighted regression was applied to separate variability in \ac{DO} associated with metabolism from tidal variation or other advection in estuaries, thereby helping to partially relax this assumption and improve estimates of ecosystem metabolism. The method was developed and tested using a simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four water quality stations within the \acl*{NERRS}.  Overall, the approach is a useful way to reduce variability in estimates of ecosystem metabolism caused by advection, particularly when the magnitude of tidal influence is high. By reducing the effects of physical transport on metabolism estimates, there may be increased potential to empirically relate metabolic rates to causal factors on times scales of several days to several weeks. Estimates of variability associated with physical advection may also be more interpretable, since convolution of physical and biological effects can be reduced, providing increased potential to resolve differences in metabolism across a mosaic of estuarine habitats.

>>>>

In aquatic ecosystems, time series of \ac{DO} have been used to compute estimates of integrated ecosystem metabolism.  Central to this open water or ``Odum'' method is the assumption that the dissolved oxygen time series is a Lagrangian specification of the flow field.  However, most \ac{DO} time series are collected at fixed locations, such that the method must assume changes in dissolved oxygen principally reflect ecosystem metabolism and that effects due to advection or mixing can be neglected.  A statistical model using weighted regression was applied to separate variability in \ac{DO} associated with metabolism from tidal variation or other advection in estuaries, thereby helping to partially relax this assumption and improve estimates of ecosystem metabolism. The method was developed and tested using a simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four water quality stations within the \acl*{NERRS}.  Overall, the approach is a useful way to reduce variability in estimates of ecosystem metabolism caused by advection, particularly when the magnitude of tidal influence is high and correlations between tidal change and solar cycling are low. By reducing the effects of physical transport on metabolism estimates, there may be increased potential to empirically relate metabolic rates to causal factors on times scales of several days to several weeks. Estimates of variability associated with physical advection may also be more interpretable, since convolution of physical and biological effects can be reduced.

### 2014-11-24 assess.Rnw
revisions that include sun and tide angle diagnostics

Multiple time series were simulated by varying the conditions in \cref{do_obs_all} ((\cref{fig:sim_ex})) to evaluate weighted regression under difference conditions. Specifically, the simulated data varied in the relative amount of noise in the measurement ($e_{pro}$, $e_{obs}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$. A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).  Half-window widths (day, hour of day, and tide height) for the weighted regressions were evaluated for each time series: time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.  The window widths were chosen based on preliminary assessments that suggested a large range in model performance was described by these values.  In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2187 weighted regressions.

The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to characterize the biological \ac{DO} time series in \cref{do_obs}. Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.  Overall, the weighted regressions produced filtered time series that were similar to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).  The median correlation between the filtered and biological values for all time series and window widths was \Sexpr{form_fun(perf_medians['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} (very poor) to \Sexpr{form_fun(perf_rngs[2, 'cor'])} (perfect).  Mean error was \Sexpr{form_fun(perf_medians['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'], 0, 0, 0)} (perfect) to \Sexpr{form_fun(perf_rngs[2, 'err'])} (very poor).  Simulations with very poor performance were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty. As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).

Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process error ($e_{pro}$), whereas increasing observation error contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).  Model performance decreased slightly with increasing tidal effects (i.e., increasing magnitude of $DO_{adv}$).  Increasing widths for day and tidal height windows contributed to improved model performance, whereas reduced performance was observed with increasing hour windows (\cref{tab:dtd_perf2,fig:err_surf2}).  Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.

@@

Half window widths of six days, one hour, and a tidal proportion of one half were used to filter the observed \ac{DO} time series.  Although the selection of window widths involves a degree of subjectivity, results from the simulations suggested that these values were appropriate for filtering \ac{DO} times series within the constraints of the analysis. Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.  Accordingly, the regression results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.  Daily metabolism estimates before and after filtering were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.  Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.  Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.  Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}. Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assumed that physical processes were the dominant sources of these values given the tidal characteristics at each site.  Finally, means and standard errors of metabolism estimates were evaluated before and after filtering to determine if annual aggregations were significantly different.

Filtering had significant effects on the correlations between water level changes, \ac{DO} time series, and daily integrated metabolism estimates (\cref{tab:cor_res}, see the link in the \hyperref[multi]{multimedia} section for graphical results of each case study).  Correlations of observed \ac{DO} time series with predicted tidal height were highly significant and positive at all sites, except Padilla Bay where increases in water level were associated with decreases in \ac{DO}  concentration.  The filtered \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after filtering likely because of the large sample size for each site (n $\approx$ 17,500). Comparison of metabolic rates to tidal changes before and after filtering produced inconsistent results (\cref{tab:cor_res}).  Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after filtering.  Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after filtering for production and respiration, although net metabolism estimates had reduced correlations.

The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (as a proportion of the total estimates, Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration. Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for respiration), and only slightly reduced for Elkhorn Slough (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for respiration).  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (\Sexpr{form_fun(elk_per[, 'Pg_Avg'],1,1,1)} \% change from the annual mean) and respiration (\Sexpr{form_fun(elk_per[, 'Rt_Avg'],1,1,1)} \%) for Elkhorn Slough, increased mean production (\Sexpr{form_fun(pdb_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(pdb_per[, 'Rt_Avg'],1,1,1)} \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Mean net ecosystem metabolism was unchanged for all sites.  Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after filtering.

@@

\subsection{Effects of aggregation and importance of filtering}

A point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.  The effects of tidal variation on daily estimates may not be relevant if seasonal or annual aggregations (e.g., mean annual metabolism) remove this potential bias.  The example from Sapelo Island in the previous section highlights this point given that mean production and respiration estimates before and after filtering were generally unchanged for the two-week period. Alternatively, annual averages of production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay (\cref{tab:case_res}). Given these results, tidal variation may or may not have effects on metabolism estimates on aggregated time scales, depending on the location.  Therefore, an evaluation of weighted regression to filter the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for its application.  Specifically, when should filtering be applied if aggregation of observed data on longer time periods removes potential bias?  A comparison of observed and filtered estimates that are aggregated over different periods of observation could help address this question.

The observed and filtered daily estimates were averaged by month for each case study to evaluate effects of aggregation on mean production and respiration.  Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with the aggregations. Significant variation in aggregated production and respiration estimates was observed for each case study (\cref{fig:metab_sum}).  Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months.  Specifically, metabolism estimates based on observed \ac{DO} were substantially muted for both Padilla Bay and Rookery Bay during summer months, whereas values were significantly higher after filtering. Results for Sapelo Island suggested that aggregated estimates were similar before and after filtering, although winter and summer months were slightly under- and over-estimated, respectively, using the observed data.  Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering.  Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.  Each case study showed differences in observed and filtered values at monthly aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).  Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression given careful consideration of appropriate window widths.

>>>>

Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under difference conditions. Specifically, the simulated data varied in the relative amount of noise in the measurement ($e_{pro}$, $e_{obs}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$. A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).  Half-window widths (day, hour of day, and tide height) for the weighted regressions were evaluated for each time series: time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.  The window widths were chosen based on preliminary assessments that suggested a large range in model performance was described by these values.  In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2187 weighted regressions.

The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to characterize the biological \ac{DO} time series in \cref{do_obs}. Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.  Overall, the weighted regressions produced filtered time series that were similar to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).  The median correlation between the filtered and biological values for all time series and window widths was \Sexpr{form_fun(perf_medians['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} (very poor) to \Sexpr{form_fun(perf_rngs[2, 'cor'])} (perfect).  Mean error was \Sexpr{form_fun(perf_medians['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'], 0, 0, 0)} (perfect) to \Sexpr{form_fun(perf_rngs[2, 'err'])} (very poor).  Simulations with very poor performance were those those with the \ac{DO} signal composed entirely of noise from observation uncertainty. As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).

Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process error ($e_{pro}$), whereas increasing observation error contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).  Model performance was not significantly affected by increasing tidal effects (i.e., increasing magnitude of $DO_{adv}$).  Model performance was not substantially affected by variation in half window widths relative to characteristics of the \ac{DO} time series (\cref{tab:dtd_perf2,fig:err_surf2}).  Graphical summaries of model performance averaged by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.

@@

Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.  The selection of half-window widths for filtering the \ac{DO} time series was based on a weight-of-evidence approach using four performance metrics to evaluate the results.  First, the regression results were evaluated using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.  Daily metabolism estimates before and after filtering were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.  Production rates were compared to mean rates of tidal height change during the day and respiration rates were compared to mean rates of change during the night.  The second and third performance metrics evaluated changes in the mean and standard error of annual metabolism estimates before and after filtering.  We assumed that mean values that represent annual aggregations would be unchanged because tidal variation is primarily diurnal, whereas the variance (i.e., standard error) would be reduced after filtering.  Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.  Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}. Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assumed that physical processes were the dominant sources of these values given the tidal characteristics at each site.

Multiple combinations of half-window widths were evaluated for the case studies.  Specifically, half-window widths of 1, 3, 6, 9, and 12 days, 1, 3, 6, 9, and 12 hours, and 0.2, 0.4, 0.6, 0.8, and 1 tidal height proportions were evaluated, producing a total of 125 unique combinations.  Half-window widths that maximized the four performance metrics for each case study were chosen as the `optimal' values.  Accordingly, the optimal half-window widths were \Sexpr{paste(casewinstxt$ELKVM)} (days, hours, tidal height) for Elkhorn Slough, \Sexpr{paste(casewinstxt$PDBBY)} for Padilla Bay, \Sexpr{paste(casewinstxt$RKBMB)} for Rookery Bay, and \Sexpr{paste(casewinstxt$SAPDC)} for Sapelo Island.  Filtering had significant effects on the correlations between water level changes, \ac{DO} time series, and daily integrated metabolism estimates (\cref{tab:cor_res}, see the link in the \hyperref[multi]{multimedia} section for graphical results of each case study).  Correlations of observed \ac{DO} time series with predicted tidal height were positive at all sites, except Padilla Bay where increases in water level were associated with decreases in \ac{DO}  concentration.  The filtered \ac{DO} time series had correlations with tidal height close to zero. Similarly, metabolic rates (production, respiration) estimated from the filtered \ac{DO} time series had significantly reduced correlations with tidal height change.

The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (as a proportion of the total estimates, Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration. Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (\Sexpr{form_fun(elk_per[, 'Pg_Avg'],1,1,1)} \% change from the annual mean) and respiration (\Sexpr{form_fun(elk_per[, 'Rt_Avg'],1,1,1)} \%) for Elkhorn Slough, increased mean production (\Sexpr{form_fun(pdb_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(pdb_per[, 'Rt_Avg'],1,1,1)} \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.  The implications of these undesirable results are described below.  Decreases in the standard erorr for all metabolism estimates were observed for all cases after filtering.

@@

\subsection{Accuracy of results and effects of aggregation}

A point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.  The effects of tidal variation on daily estimates may not be of concern if seasonal or annual aggregations (e.g., mean annual metabolism) remove this potential bias.  The example from Sapelo Island in the previous section highlights this point given that mean production and respiration estimates before and after filtering were generally unchanged for the two-week period. Alternatively, annual averages of production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay but not Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Therefore, an evaluation of weighted regression to filter the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for its application.  Specifically, does the period of observation affect the ability of weighted regression to remove physical variation in the time series?  When should filtering be applied if aggregation of observed data on longer time periods removes potential bias?  The first question is addressed by evaluating collinearity between tidal change and solar periods.  The second question is addressed by comparing observed and filtered estimates that are aggregated over different periods of observation.

Collinearity between tidal height change and solar cycling likely affected the ability of weighted regression to quantify the variation in \ac{DO} time series.  Model parameterization may be unreliable if, for example, tidal height follows diurnal periods by increasing during the day or decreasing during the night.  \citet{Nidzieko14} found that such covariation is common in Elkhorn Slough during the summer months when high tides always occurred during the night.  Given that the phasing between tidal height change and diurnal cycling is variable, the ability of weighted regression to quantify variation attributed to both is also expected to vary.   The correlation between sun angle and tidal change (measured as an angular rate) was evaluated using a moving window approach for each case study.  This approach is analogous to weighted regression by providing an indication of when collinearity may occur as a function of the moving window.  Weighted regression can be expected to effectively characterize biological and physical variation during periods when the correlation between tidal height change and diurnal cycling is low within the window.  \Cref{fig:move_corrs} suggests that collinearity between sun angle and tidal height change can exceed +/-0.2 for Elkhorn Slough and Padilla Bay, whereas correlations were much smaller regardless of the period of observation for Rookery Bay and Sapelo Island.  Given the change in mean annual metabolism using the filtered \ac{DO} time series and the relatively high collinearity between tidal change and solar cycling, the results for Elkhorn Slough and Padilla Bay may not be accurate.  The results may only be interpretable when correlations are close to zero (e.g., April and October for Elkhorn Slough).

The observed and filtered daily estimates were averaged by month for each case study to evaluate effects of aggregation on mean production and respiration, in addition to the mean annual estimates in \cref{tab:case_res}. Significant variation in aggregated production and respiration estimates was observed for each case study (\cref{fig:metab_sum}).  Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months.  Specifically, metabolism estimates based on observed \ac{DO} were substantially muted for both Padilla Bay and Rookery Bay during summer months, whereas values were significantly higher after filtering. Results for Sapelo Island suggested that aggregated estimates were similar before and after filtering, although winter and summer months were slightly under- and over-estimated, respectively, using the observed data.  Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering, which may have been related to collinearity.  Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.  Each case study showed differences in observed and filtered values at monthly aggregations, suggesting tidal variation may influence metabolism estimates at relatively long time scales \cref{tab:case_res}).

### 2014-11-24 comments.Rnw
final proof after MCM revisions
revisions that include sun and tide angle diagnostics

Results from the simulations and case studies suggested that weighted regression can be a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.  However, application of the method may only be appropriate under specific situations.  The case studies were chosen based on the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.  Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.  The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.  Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.  In particular, alternative window widths should evaluated for the ability to remove tidal effects while preserving the biological signal.  The window widths in the above analysis may have removed variation in the \ac{DO} signal from both of these sources.

Although the above analyses suggest the approach has merit, the case studies emphasize a critical challenge in applying weighted regression to monitoring data. Specifically, the true biological signal is not known and the relative contribution of horizontal advection to bias is not accurately quantified with the available data.  Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate the performance of weighted regression.  In the absence of additional validation, we propose a precautionary approach for application of the weighted regression to monitoring data.  Weighted regression may be most effective at macrotidal sites with strong evidence of the effects of tidal advection on biological signals.  A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest detiding is appropriate.  The choice of window widths may also produce varying results.  Window widths that produce large changes in mean annual estimates should be interpreted with caution.  In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.  Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.

>>>>

Results from the simulations and case studies suggested that weighted regression can be a practical approach for filtering \ac{DO} time series.  However, application may only be appropriate under specific situations.  The case studies were chosen based on the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.  Despite these similarites, filtering had variable effects on metabolism estimates.  The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.  Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, collinearity between tidal change and solar cycling likely contributed to the relative change in mean metabolism before and after filtering.  Evaluating collinearity, as in \cref{fig:move_corrs}, is an important diagnostic for characterizing the expected performance of weighted regression and choosing appropriate window widths.  These plots should indicate whether weighted regression is appropriate for the whole time series (e.g., Sapelo Island) or only for specific periods when collinearity is low (e.g., April and October at Elkhorn Slough).  A weight-of-evidence approach should also be used such that application of weighted regression should reduce anomalous metabolism estimates, minimize correlations between observed \ac{DO} and tide height, and reduce clear visual patterns of tide change on \ac{DO}.  In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.  Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.

### 2014-11-24 dep_dat.Rnw
some edits from Jim's comments, windows specific to each case study
revisions that include sun and tide angle diagnostics

@@

######
# changes in mean production, respiration for each site after filtering
# used in estimates of ecosytem metab before/after section in assessment
# basically same code used to create table 5
load('met_comp.RData')

@@

######
# objects for selecting window widht combinations for case studies
casewins <- list(6, 1, 0.5)
@

>>>>

# objects for selecting window widht combinations for case studies
casewins <- list(
  ELKVM = list(12, 6, 0.8), 
  PDBBY = list(3, 6, 0.6), 
  RKBMB = list(3, 1, 0.6), 
  SAPDC = list(3, 1, 0.6)
)

# casewins as text
casewinstxt <- llply(casewins, function(x){
  tmp <- unlist(x)
  tmp[3] <- paste('and', tmp[3])
  paste(tmp, collapse = ', ')
})

# subset metab estimates by window widths for each case study
load('case_grds.RData')
sel_vec <- llply(
  casewins, 
  .fun = function(x) with(case_grds, which(dec_time == x[[1]] & hour == x[[2]] & Tide == x[[3]]))
)
sel_vec <- paste0(names(casewins), '_wtreg_', unlist(sel_vec), '.RData')

met_comp <- met_comp[met_comp$.id %in% sel_vec, ]

@@

######
# changes in mean production, respiration for each site after filtering
# used in estimates of ecosytem metab before/after section in assessment
# basically same code used to create table 5

@@

@

### 2014-11-24 discussion.Rnw
revisions that include sun and tide angle diagnostics

The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.  The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove variation in observed \ac{DO} from tidal change.  Further, aggregation of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location. These results suggest that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.  Additionally, variation of aggregated metabolism estimates were substantially reduced after filtering, suggesting greater confidence in interpreting estimates even if the mean values are similar.

@@

Correlations of metabolism estimates with tidal height changes after filtering were generally reduced, although trends were not always consistent.  However, correlations of net metabolism estimates were reduced in all cases.  An additional indication of the effectivenes of weighted regression was the reduction of anomalous metabolism estimates after filtering for all case studies.  Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, although `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.  For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime such that inundation of salt marshes during the night following by draining with low tide during the day lead to inflated respiration values.   Synchrony between solar and tidal cycles is a critical concern for interpreting metabolism estimates, although a broader discussion regarding whether or not this represents an actual bias in metabolism from physical advection may be needed.

>>>>

The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.  The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove variation in observed \ac{DO} from tidal change.  Further, aggregation of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location. These results suggest that previous estimates may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.  Additionally, variation of aggregated metabolism estimates were substantially reduced after filtering, suggesting greater confidence in interpreting estimates even if the mean values are similar.

@@

For all case studies, weighted regression was generally successful in reducing the variation in the \ac{DO} time series that was caused by physical advecttion.  In particular, the reduction of anomalous metabolism estimates after filtering was observed.  Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, although `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.  For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime such that inundation of salt marshes during the night following by draining with low tide during the day lead to inflated respiration values.  Synchrony between solar and tidal cycles is a critical concern for interpreting metabolism estimates such that collinearity between the two may diminish the performance of weighted regression.
